Model to study co-evolution of trade an culture

### 2015-03-31 article.tex
new result and Trade Model

In order to validate our model a result from the literature on cultural transmission is reproduced. We then illustrate the possibility to implement a trade mechanisms.

@@

In this article the neutral hypothesis is tested which implies that the cultural transmission mechanism is the ``random copy"~\cite{bentley_random_2004,bentley_specialisation_2005,mesoudi_random_2009}. The aim is to show that our model is indeed able to reproduce the main trends expected with such copy mechanism. Namely, we should observe a distribution known as the ``power law'', displaying a line in the log-log graph of the proportion of variants with regards to the frequency of variant.

@@

The score reflects the ability for an agent to obtain the products it needs. It is maximum when the quantity vector of an agent is equal to its need vector and lower in proportion to the distance between the need vector and the quantity vector.  It is formally computed as follow for agent $i$:

$$ 1- \dfrac{\abs{q^i_j - n_j}}{ \sqrt{\abs{(q^i_j)^2-(n_j)^2}}} $$

This function ensures that each good count for the same proportion in the final score (i.e.: managing to get only the right amount of a good with a high ``need'' value will not give a better score to the agent).

@@

\todo{miss the formula for a possible trade}

@@

$$ hey $$

@@

With such a setup it is expected that the prices will converge to a ratio allowing each agent to obtain a number of resources equivalent to the needs. Knowing that,  the best possible price of a good should follow the equation :

$$\frac{V^j_i}{V^j_g} = N_i   $$
with g the produced good of the agent j and i another good which is not the one which is produced.

If such a price is reached to score of all agents should be equal to : $$ 10\times1\times(ngoods-1) $$

@@

\begin{figure}[hbp]
	\begin{center}
%		\includegraphics[width=7cm]{img/allmuRandMaxBentleyValues.pdf}
	\end{center}
	\caption{Distribution of frequencies depending on the $\mu$ parameter}
	\label{fig:allMutation}
\end{figure}

@@

\begin{table}[h]
	\centering
	\begin{tabular}{cc}
		\begin{tabular}{l|cc}
			$\mu$ & $\alpha$ & SD\\\hline
			0.004&1.541193&0.031384174\\
			0.008&1.549726&0.021953251\\
			0.016&1.568096&0.017025456\\
			0.064&1.651690&0.006538209\\

		\end{tabular}
		&
		\begin{tabular}{l|cc}
			$\mu$ & $\alpha$ & SD\\\hline
			0.004&1.539710&0.022515901\\
			0.008&1.557460&0.017805810\\
			0.016&1.570036&0.004412181\\
			0.064&1.651521&0.009988488\\
		\end{tabular}\\
		(a) & (b)
	\end{tabular}
	\caption{Mean \& SD are done with 100 run for each with 250 agent for (a) and 500 for (b) }
	\label{tab:mualpha}
\end{table}

@@

\simon{What do you think about puting  the bentley et al resutlts cf table~\ref{tab:Bmualpha} below}

\begin{table}
	\centering
	\begin{tabular}{l|cc}
N=250\\\hline
$\mu$&$\alpha$&(SD)\\\hline
0.004&1.50&(0.02) \\
0.008&1.52&(0.02)\\
0.016&1.55&(0.03)\\
0.032&1.59&(0.06)\\
0.064&1.78&(0.08)\\
0.128&1.99&(0.04)\\
	\end{tabular}
	\begin{tabular}{l|cc}
N=500\\\hline
$\mu$&$\alpha$&(SD)\\\hline
0.004&1.53&(0.03)
0.008&1.57&(0.01)
0.016&1.61&(0.04)
0.032&1.70&(0.05)
0.064&1.81&(0.10)
0.128&1.98&(0.02)
	\end{tabular}
	\caption{Mean \& SD from \cite{bentley_random_2004} }
	\label{tab:Bmualpha}
\end{table}

@@

\todo{add innovation and imitation}
\begin{table}[h]
	\centering
	\begin{tabular}{l|c|c}
	 Imitation  Process :& Rand & Eco \\
	 Mutation Process& & \\\hline
		$\mu$ &A & B \\
		$\mu +\alpha$ & C & D \\
	\end{tabular}
	\caption{Experimental Setup}
	\label{tab:exp}
\end{table}

@@

\begin{figure}[H]
	\begin{center}
		\includegraphics[width=7cm]{img/4SetupDistrib.png}
	\end{center}
	\caption{Frequency distribution for the 4 setup}
	\label{fig:4setDi}
\end{figure}

@@

\begin{table}[H]
	\centering
	\begin{tabular}{l c c}
		& Random Transmission & Eco Transmission \\
		$\mu$ & \specialcell{\raisebox{-.5\height}{\includegraphics[width=5cm]{img/scoreWrtTimeA.png}}\\Setup A}&\specialcell{\raisebox{-.5\height}{\includegraphics[width=5cm]{img/scoreWrtTimeC.png}} \\ Setup B} \\
		$\mu + \alpha$& \specialcell{\raisebox{-.5\height}{\includegraphics[width=5cm]{img/scoreWrtTimeB.png}} \\ Setup C}&\specialcell{\raisebox{-.5\height}{\includegraphics[width=5cm]{img/scoreWrtTimeD.png}} \\ Setup D}

	\end{tabular}
	\caption{Evolution of the score in the different setup}
	\label{tab:scoreEvol}
\end{table}

>>>>

In order to validate our model we first reproduced traditional  result often seen in the literature on cultural transmission. After that we show that doing small modification we can transform our model to make it fit processes that are more economically consistent.

@@

In this article the neutral hypothesis is tested which implies that the cultural transmission mechanism is the ``random copy"~\cite{bentley_random_2004,bentley_specialisation_2005,mesoudi_random_2009}. In the literature, people have shown that a lot of cultural phenomene where people copy randomly the cultural traits of other exhibit a particular pattern.

With this mechanism of copy and given enough time  we will find in a fixed size population always the same repartition of the cultural variants. A few that are very frequent and an always bigger number of variant that are always less frequent. This is traduced by a ``power law'' distribution, following the equation $P(v)=C/v^\alpha $ that one could display as a line with a slope $-\alpha$ in the log-log graph of the proportion of variants with regards to the frequency of variant.

@@

The score $S^i$ of agent $i$ reflects the ability for an agent to obtain the products it needs. It is maximum when the quantity vector of an agent is equal to its need vector and lower in proportion to the distance between the need vector and the quantity vector.  It is formally computed as follow for agent $i$:

$$ S^i=1- \dfrac{\abs{q^i_j - n_j}}{ \sqrt{\abs{(q^i_j)^2-(n_j)^2}}} $$

This function ensures that each good count for the same proportion in the final score (i.e.: managing to get only the right amount of a good with a high ``need'' value will not give a better score to the agent). This can be seen in the figure~\ref{fig:fit}.

@@

@@

Let be : $o$ the agent ``offerer'', and $r$ the agent that receive the offer. Both have an inventory $x_i$ of every good where $x_i^o$ is the quantity of good $i$ owned by $o$. $o$ produce the good $g$ and want to trade it in a quantity $q_g^o$ for a quantity $q_k^o$ of the good $k$ produced by $r$. 
In its side $r$ own also a quantity $q_k^r$ of the good $g$ that it want to trade with quantity $q_k^r$ of its own produced goods.

The exchange take place only if :
$$ x_g^o >= q_g^o $$
$$ x_g^r <= q_g^o $$
$$ x_k^r >= q_k^o $$
$$q_g^o>=(x_g^r+q_g^r)$$
$$q_k^o<=q_k^r$$
$$q_g^o<=q_g^r$$

The given quantity of the produced good ($g$ for $o$) is the estimate value of it ( its price). The requested quantity of the non produced value is simply the ratio between the estimated value of the good requested and the given quantity of the produced good.

In our case, for $o$ : $q_g^o = v_g^o$ and $q_k^o= \frac{v_k^o}{v_g^o} $ and for $r$  : $q_k^o = v_k^o$ and $q_g^o= \frac{v_g^o}{v_k^o} $.

@@

With such a setup it is expected that the prices will converge to a ratio allowing each agent to obtain a number of resources equivalent to the needs. Knowing that,  the best possible price of a good should satisfy the equations :

$$\frac{v^o_k}{v^o_g} = n_k $$
$$\frac{v^r_g}{v^r_k} = n_g $$
$$v^r_k = \frac{1}{n_k} $$
$$v^r_g = \frac{1}{n_g} $$
with g the produced good of the agent j and i another good which is not the one which is produced.

If such a price is reached to score of all agents should be equal to : $$ CulturalStep\times1\times ngoods $$

@@

\begin{figure}[hbp]
	\begin{center}
		\includegraphics[width=7cm]{img/allmuRandMaxN250.pdf}
		\includegraphics[width=7cm]{img/allmuRandMaxN500.pdf}
	\end{center}
	\caption{Distribution of frequencies depending on the $\mu$ parameter with 250 agents (left) and 500 agents (right)}
	\label{fig:allMutation }
\end{figure}

@@

\begin{table}[h]
	\centering
	\begin{tabular}{l|cc}
		Agents Number & Our result & Bentley et al 2004\\\hline
		N = 250 &
		\begin{tabular}{l|cc}
			$\mu$ & $\alpha$ & SD\\\hline
			0.004&1.53&0.03\\
			0.008&1.55&0.03\\
			0.016&1.57&0.02\\
			0.064&1.66&0.01\\
		\end{tabular}
		&
		\begin{tabular}{cc}
		$\alpha$&SD\\\hline
		1.50&0.02 \\
		1.52&0.02\\
		1.55&0.03\\
		1.78&0.08\\
		\end{tabular} \\
		N= 500&
		\begin{tabular}{l|cc}
			$\mu$ & $\alpha$ & SD\\\hline
			0.004&1.54&0.02\\
			0.008&1.55&0.01\\
			0.016&1.57&0.01\\
			0.064&1.67&0.01\\
		\end{tabular}
		&
		\begin{tabular}{cc}
			$\alpha$&(SD)\\\hline
			1.53&0.03\\
			1.57&0.01\\
			1.61&0.04\\
			1.81&0.10\\
		\end{tabular}
	\end{tabular}
	\caption{Mean \& SD are calculate on 100 run.}
	\label{tab:mualpha}
\end{table}

@@

\simon{What do you think about puting  the bentley et al resutlts cf table~\ref{tab:Bmualpha} below?\\ 
By the way, doing t.test on the different value of alpha I have gives me statistically differences between 250 and 500 agents ONLY for $\mu=.064$  }

@@

\todo{add innovation and imitation}
\begin{table}[h]
	\centering
	\begin{tabular}{l|cc}
					&Imitation  Process \\
		Inovation Mechanism	& Rand & Success \\\hline  
		Unbounded 		&A & B \\
		Local Modification	& C & D \\
	\end{tabular}
	\caption{Experimental Setup}
	\label{tab:exp}
\end{table}

@@

\begin{figure}[H]
	\begin{center}
		\includegraphics[width=7cm]{img/frequenciesABCD.pdf}
	\end{center}
	\caption{Frequency distribution for the 4 setup}
	\label{fig:4setDi}
\end{figure}

\begin{table}[h]
	\centering
	\begin{tabular}{lcccc}
		& A & B & C & D\\
A & X& \includegraphics[width=5cm]{img/A-B.pdf}&\includegraphics[width=5cm]{img/A-C.pdf}&\includegraphics[width=5cm]{img/A-D.pdf}\\
B & X& X &\includegraphics[width=5cm]{img/B-C.pdf}&\includegraphics[width=5cm]{img/B-D.pdf}\\
C & X& X & X &\includegraphics[width=5cm]{img/C-D.pdf}\\
	\end{tabular}
	\caption{Experimental Setup}
	\label{tab:allRes}
\end{table}

@@

\begin{table}[H]
	\centering
	\begin{tabular}{l c c}
		& Random Transmission & Eco Transmission \\
		$\mu$ & \specialcell{\raisebox{-.5\height}{\includegraphics[width=5cm]{img/ScoreEvolutionForA-G2N250.pdf}}\\Setup A}&\specialcell{\raisebox{-.5\height}{\includegraphics[width=5cm]{img/ScoreEvolutionForB-G2N250.pdf}} \\ Setup B} \\
		$\mu + \alpha$& \specialcell{\raisebox{-.5\height}{\includegraphics[width=5cm]{img/ScoreEvolutionForC-G2N250.pdf}} \\ Setup C}&\specialcell{\raisebox{-.5\height}{\includegraphics[width=5cm]{img/ScoreEvolutionForD-G2N250.pdf}} \\ Setup D}

	\end{tabular}
	\caption{Evolution of the score in the different setup for typical run of each setup.}%%
	\label{tab:scoreEvol}
\end{table}
